Variational Monte Carlo on bosonic systems

### 2018-03-01 project_1/article/article.tex
Blocking

	\subsection{Statistical Analysis}
		If the results of the metropolis sampling were completely uncorrelated,
		it would be enough to compute the standard deviation in a familiar way,
		\begin{equation}
			\sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
			\label{eq:standard_deviation}
		\end{equation}
		where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
		However, it is reasonable to assume that the data we are dealing with in this
		study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
		does not hold. The prevailing
		definition of autocorrelation in a data stream or signal is correlation 
		between a delay of the signal and the original signal. One would be interested to find the
		delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
		spacing as $d$, and define the following correlation function,
		\begin{equation}
			f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
			\label{eq:correlation_f_d}
		\end{equation}
		The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
		would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
		\begin{equation}
			\kappa_d = \frac{f_d}{\text{Var}(x)},
			\label{eq:autocorrelation_function}
		\end{equation}
		which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
		function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
		\begin{equation}
			\tau = 1 + 2 \sum_{d=1}^{n-1},
		\end{equation}
		notice that the autocorrelation time is $1$ for a correlation free experiment.
\end{document}

>>>>

	\subsection{Statistical Analysis}
		If the results of the metropolis sampling were completely uncorrelated,
		it would be enough to compute the standard deviation in a familiar way,
		\begin{equation}
			\sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
			\label{eq:standard_deviation}
		\end{equation}
		where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
		However, it is reasonable to assume that the data we are dealing with in this
		study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
		does not hold. The prevailing
		definition of autocorrelation in a data stream or signal is correlation 
		between a delay of the signal and the original signal. One would be interested to find the
		delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
		spacing as $d$, and define the following correlation function,
		\begin{equation}
			f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
			\label{eq:correlation_f_d}
		\end{equation}
		The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
		would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
		\begin{equation}
			\kappa_d = \frac{f_d}{\text{Var}(x)},
			\label{eq:autocorrelation_function}
		\end{equation}
		which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
		function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
		\begin{equation}
			\tau = 1 + 2 \sum_{d=1}^{n-1},
		\end{equation}
		notice that the autocorrelation time is $1$ for a correlation free experiment.

		We are now able to make a correction to the expression for the standard deviation 
		improving on \autoref{eq:standard_deviation} by taking correlation into account,
		\begin{equation}
			\sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}(\ev{E^2_L} - \ev{E_L}^2)},
		\end{equation}
		where $\Delta t$ is the time between each sample. The main problem at this point is to 
		find $\tau$, which we do not know for any given system and it is generally very expensive to
		compute. In order to find a good estimate of $\tau$ we use a procedure called \emph{blocking}.

		\subsubsection{Blocking}
		In the method of blocking we group the samples into blocks of increasing size. If one where to compute
		the standard deviation for each block, one should see the variance increase with the block size.
		The standard deviation would only increase up to a certain point, from whence it would stay
		almost constant. What is happening is that we have reached a point where a particular
		sample from one block is no longer correlated with a corresponding sample from an adjacent block.
		The block size for this point of convergence now functions as an estimate for the autocorrelation 
		time $\tau$.

\end{document}
